One-dimensional turbulence (ODT): computationally efficient modeling and simulation of turbulent flows

### 2020-04-29 manuscript/manuscript.tex
Additions to motivation section. Lots of references added. Added commented text for example section.

@@

@@

%Example cases:
%\begin{itemize}
%	\item From cylindrical ODT paper:
%	\begin{enumerate}
%		\item pipe flow
%		\item nonreacting round jet
%		\item round jet flame
%	\end{enumerate}
%	\item Other possible example cases:
%	\begin{itemize}
%		\item something to compare with DNS/LES results to illustrate efficiency
%		\item no sooting cases since no soot in the code
%		\item planar vs. cylindrical comparison
%	\end{itemize}
%\end{itemize} 
%
%Things to talk about with example cases:
%\begin{itemize}
%	\item compare to DNS/LES cases; computational efficiency, accuracy
%	\item how data looks before post processing, how post processing works
%\end{itemize}

@@

%COPIED/PASTED FROM CYLINDRICAL ODT PAPER
%ODT has been applied by a number of researchers to a wide range of flows. These include homogeneous turbulence \cite{Kerstein_1999,Sun_2014}, wakes \cite{Kerstein_2000}, mixing layers \cite{Kerstein_1999,Kerstein_2001,Ashurst_2003,Ashurst_2005}, stratified flows \cite{Wunsch_2001}, Rayleigh-Benard flows \cite{Wunsch_2005}, buoyant wall flows \cite{Dreeben_2000,Shihn_2007}, and channel flows \cite{Schmidt_2009,Schmidt_2003,Lignell_2013}. Schmidt et al. \cite{Schmidt_2013} studied buoyantly-driven cloud-top entrainment with comparisons to water tank experiments. Gonzalez-Juez et al. \cite{Esteban_2013} studied reactive Rayleigh-Taylor mixing. A number of ODT simulations of turbulent jet flames have been performed \cite{Echekki_2001,Hewson_2001,Hewson_2002,Punati_2011,Lignell_2012b,Lignell_2015}. Other combustion applications include wall fires \cite{Monson_2016,Shihn_2004}, buoyant pool fires \cite{Ricks_2010,Hewson_2006,Hewson_2009}, and opposed jet flames \cite{Jozefik_2015}. Several studies used DNS to validate ODT for combustion in temporal planar jets \cite{Punati_2011,Lignell_2012b,Lignell_2015}. ODT has also been applied as a subgrid model for large eddy simulations (LES) \cite{Cao_2008,Schmidt_2003,Schmidt_2010} and in particle dynamics simulations in channel flows \cite{Schmidt_2009}, homogeneous turbulence \cite{Sun_2014}, and jets \cite{Sun_2017,Goshayeshi_2015}.

%TEXT COPIED FROM ABANDONED JOSS DRAFT

%ODT has been applied to a wide range of flows. Early applications focused on homogenous turbulence, wakes, and mixing layers (Kerstein1999,Kerstein2000,Kerstein2001). Later extension to variable-density flows and a spatial downstream coordinate system facilitated its growth and application to more complex flows, including combustion in jet flames (Echekki2001; Hewson2001; Hewson2002; Lignell2012; Punati2011; Abdelsamie2017; Lignell2017; Goshayeshi2015), counterflow flames  (Jozefik2015), wall fires (Monson2016), and sooting flames, (Lignell2015; Hewson2006; Hewson2009; Lignell2015b, Ricks2010), as well as other particle flows (Sun2017; Schmidt2009; Sun2014; Fistler2017).  ODT has also served to complement LES through subgrid modeling studies (Cao2008; Schmidt2003; Schmidt2010) and has been applied to various other flow configurations such as double-diffusive interfaces (GonzalezJuez2011), Rayleigh-Taylor mixing (GonzalezJuez2013), and stratified turbulence (Wunsch2001).  Most recently, the ODT code was extended to include cylindrical and spherical coordinate systems (Lignell2018; Klein2018; Klein2019). During this implementation, the ODT code was drastically overhauled, resulting in the SEC package presented here. Further detail on the current ODT model formulation and implementation can be found in the literature (Ashurst2005; Kerstein2001; Lignell2013; Lignell2018).

%Ongoing research involving ODT and SEC spans multiple research groups and subject areas. ODT is currently being used to investigate soot formation and destruction in ethylene jet flames via parametric modeling studies (Hewson2015; Lansinger2015b). The SEC package framework is also being used to develop the hierarchical parcel swapping model (HiPS), which models turbulence with an economical, hierarchical network that represents the fluid as individual parcels capable of switching positions within the tree (Kerstein2013; Kerstein2014). In the future, SEC may also be extended to include the linear eddy model (LEM) (Kerstein1991) and LES approaches.

>>>>

Early applications of ODT focused on homogenous turbulence, wakes, and mixing layers \cite{Kerstein_1999,Kerstein_2000,Kerstein_2001}. Later extension to variable-density flows and a spatial downstream coordinate system facilitated its growth and application to more complex flows, including combustion in jet flames \cite{Echekki_2001,Hewson_2001,Hewson_2002,Lignell_2012,Punati_2011,Abdelsamie_2017,Lignell_2017, Goshayeshi_2015}, counterflow flames \cite{Jozefik_2015}, wall fires \cite{Monson_2016}, and sooting flames \cite{Lignell_2015,Hewson_2006,Hewson_2009,Lignell_2015b,Ricks_2010}, as well as other particle flows \cite{Sun_2017,Schmidt_2009,Sun_2014,Fistler_2017}. ODT has also served to complement LES through subgrid modeling studies \cite{Cao_2008,Schmidt_2003,Schmidt_2010} and has been applied to various other flow configurations such as double-diffusive interfaces \cite{GonzalezJuez_2011}, Rayleigh-Taylor mixing \cite{GonzalezJuez_2013}, and stratified turbulence \cite{Wunsch_2001}. Most recently, the ODT code was extended to include cylindrical and spherical coordinate systems \cite{Lignell_2018,Klein_2018,Klein_2019}.

During the implementation of the cylindrical and spherical model formulations, the ODT code was drastically overhauled and reorganized, resulting in the current development version of the code. The ODT code presented here is a pared down version of the development code, representing the most fundamental aspects of the ODT model. To this end, the code presented here does not include all of the development code's functionalities. The example cases in Section \ref{sec:examples} represent a good sampling of the ODT code's capabilities as it is presented here.

@@

\subsection{Pipe Flow}
\label{sub:pipeflow}

% TEXT COPIED FROM CYLINDRICAL ODT PAPER
%We present results for incompressible pipe flow simulations using the temporal, cylindrical ODT formulation. Results for three different friction Reynolds numbers Reτ = 550, 1000, 2000 are compared to DNS results f Kh t l [26](R 550 1000) d Chi t l [6](R 2000) Silti lt from Khoury et al. [26](Reτ = 550, 1000) and Chin et al. [6](Reτ = 2000). Simulation results were dd i i dit f D 2 0 d fl dit f 1 0k −3 F fll d ld i fl it produced using a pipe diameter of D = 2.0 m and flow density of 1.0kg m−3. For fully developed pipe flow, it is possible to estimate the value of a constant mean pressure gradient driving the flow based on the value of the friction velocity, the pipe radius and the density. Friction velocity values of 1 (Reτ = 550, 1000) and 2 ms−1 (R 2000) d d d t llt th dit di i th fl Silti (Reτ = 2000) were assumed and used to calculate the mean pressure gradient driving the flow. Simulation results achieving statistical convergence for the friction velocity were verified afterward as a check on the input parameters. The simulations used initial conditions with constant velocity profiles. The simulations were run to a developed flow state, after which simulation data were gathered until statistical convergence for the root mean square (RMS) velocity difference from the mean profiles occurred. The total normalized run time trun/τpipe = trunu/D was 20200, 25070, and 28140 for Reτ = 550, 1000, and 2000, respectively, where u is the average velocity and D is the pipe diameter.

%The simulations were performed with parameters ofC = 5and Z = 350 for the temporalODTformulation. Additionally, a restriction was imposed on the eddy size range by selecting eddies only up to a maximum normalized size of Le,max/D = 1/3. This restriction limits the eddy size by construction, as opposed to the large-eddy suppression mechanism commonly used in ODT simulations [27,45]. The values of C, Z,and Le,max/D were adjusted to give good agreement of the ODT results compared to the DNS. Schmidt et al. [45]. showed that higher Z results in the buffer-layer being located further from the wall; increasing C results in a lower slope of the mean streamwise velocity in the log-layer; and higher Le,max/D gives a smaller mean streamwise velocity in the wake region.

%Figure 6a shows results of the mean streamwise velocity profiles in wall units for each of the three Reτ considered. The profiles are shifted vertically by 10 and 20 units in the figure for presentation. The Reτ = 550 case shows a comparison of the TMA, TMB, and the hybrid planar/cylindrical approach of the TMB eddy event implementation (PTMB). As seen in the figure, the differences between TMA, TMB, and PTMB are negligible, and so were not considered for the higher Reτ cases. The agreement between the ODT and DNS for the mean velocity is excellent for all three Reτ values.

%Figure 6b shows the streamwise root mean square (RMS) velocity profiles for each Reτ . These profiles are shifted vertically by 2 and 4 units in the figure for presentation. The ODT RMS velocity profiles deviate from the DNS more strongly than the mean velocity profiles, with the ODT value at the peak approximately 20% lower than the DNS. The qualitative shape of the profiles, however, is the same as expected from previous ODT channel flow simulations performed with the planar formulation [33,37]. The small double peak structure of the ODTwas described in [33] and arises from alignment of the triplet map images in the near-wall region. The slight depression between the ODT peaks aligns with the DNS peak, resulting in a larger difference between the RMS profiles than an extrapolation of the surrounding ODT profiles to the peak region would give. As for the mean profiles, the differences between TMA, TMB, and PTMB for the RMS velocity profiles are negligible. Small differences appear only for the RMS profiles near the centerline. We expect this because of the eddy timescale behavior seen in Fig. 5.

\subsection{Non-reacting Jet }
\label{sub:nonreactingjet}

% TEXT COPIED FROM CYLINDRICAL ODT PAPER
%The new cylindrical ODT formulation is demonstrated in a nonreacting round turbulent jet. Results are compared to the experimental data of Hussein et al. [18]. The jet consists of air issuing into air through a 1 in (0.0254 m) diameter duct. The jet exit velocity is 56.2 ms−1 and is well-approximated by a top-hat profile. The reported Reynolds number is 95,500, where Re = Dv0/ν,and D is the jet exit diameter, v0 is the jet exit velocity, and ν is the kinematic viscosity. The ODT simulations use the same diameter and velocity, but a kinematic viscosity of 1.534×10−5 m2 s−1, giving a Reynolds number of 93,056. The initial velocity profile i th ODT ilti i difid t ht fil i hih h bli ttfti f idth δ 0 1D in the ODT simulations is a modified top-hat profile in which a hyperbolic tangent function ofwidth δ = 0.1D is used on either side of the jet to smooth the transition between the jet and the free stream. Here, xc1 and xc2 are the center locations of the tanh transition, −D/2and D/2, respectively. In the spatial formulation ofODT, the streamwise velocity must be positive everywhere on the line due to v in the denominator on the right hand side of the evolution equations (see Sect. 2.4.1). As such, a small vmin = 0.1ms−1 is added uniformly to the velocity profile. v is taken to be the jet velocity of 56.2 ms−1.

%ODT simulations were performed using the TMB triplet map with parameters C = 5.25, βles = 3.5, and Z=400. The value of Z is the same as the spatial simulations in [38] and was not adjusted. The values ofC and βles were adjusted to give good agreement of the jet evolution with the experimental data. Note the very close agreement of the C and Z parameters here to the optimal values used for the pipe flow simulations where C = 5and Z = 350. This illustrates a level of robustness in the ODT parameters between the two configurations and suggests that intermediate values could be successfully applied in both configurations. The pipe flow is sensitive to Z, as noted above, but the jet is not, so Z = 350 would be preferred. Figure 7 shows results of the simulations.

%Here, 1024 independent ODT realizations were performed and results were ensemble averaged. All quantities are normalized consistent with jet similarity scaling. Downstream locations are normalized by the jet diameter D, and radial locations are normalized by (y − y0)where y is the downstream location and y0 = 4D is the virtual origin used in [18]. In the figure, v0 is the jet exit velocity and vcL is the local mean axial centerline velocity. Here, r is used to denote both the experimental radial location and the ODT line position x.

%Figure 7ashows v0/vcL versus y/D; the similarity scaling gives a nominally linear profile where v decays as 1/y. The ODT simulation compares very well with the stationary wire data in [18] after an initial induction period for y/D < 20. The dashed line in the plot is the linear curve fit reported by Hussein et al. Figure 7b shows radial profiles of the mean axial velocity normalized by the local centerline value. Profiles at three axial locations, 50D,70D, and 90D, are shown. The experimental data points shown are the laser Doppler some spread in the profiles at higher r/(y − y0), with the simulation results relaxing toward the experimental values with downstream distance. Figure 7c shows radial profiles of the axial root mean square (RMS) velocity anemometer (LDA) data in [18]. TheODT results show a similarity collapse of the data at r/(y−y0)< 0.1, but normalized by the local centerline velocity at the same three positions. Here again, the profiles tend to relax toward the experimental (LDA) data at higher r/(y − y0) with downstream distance.

%the quantitative agreement of vrms velocity is generally good, especially for r/(y − y0)> 0.03. At lower values, near the centerline, the ODT vrms increases, whereas the experimental data decrease slightly. We attribute this to the so-called centerline anomaly of the cylindrical ODT formulation due to the geometric stretching associated with the cylindrical triplet maps. This stretching is shown in Figs. 2 and 3.Infact,the motivation for developing TMB was to minimize the degree of this geometric stretching. The stretching effect was present for both maps TMA and TMB, but it is largest near the centerline, where curvature is large. With increasing distance from the centerline, that is, for large distances compared to the eddy size, the geometric stretching effect becomes negligible for both maps, and they both approach the planar limit.

%For comparison to the ODT results with TMB shown in Fig. 7, similar simulations were performed with TMA, and corresponding results are shown in Fig. 8. Here, βles and Z are the same as for the previous case with TMB, but C = 7. Again, we see good agreement for the centerline velocity decay in Fig. 8a. A somewhat better similarity collapse than for TMB occurs in the radial velocity profiles, though they are slightly lower than the experimental data at r/(y − y0)< 0.13 and slightly higher than the experimental data at r/(y − y0)> 0.13. The vrms profiles also show strong similarity scaling. Compared with TMB, however, the simulations show a much higher centerline anomaly, and the vrms profile departs from the experiments much further from the centerline and has a higher rise in the region r/(y − y0)< 0.05.

%It is worth noting that the vrms profile depends somewhat on the ODT parameters selected. Hewson and Kerstein [15] demonstrated that mean centerline values of mixture fraction were insensitive to C and βles provided their ratio was kept constant. In these simulations, however, higher values of RMS fluctuations were observed at lower values of βles. Figure 9 shows results for TMA similar to those in Fig. 8 with varying C and βles. Each of the three cases has C/βles = 2, with C = 7, 5, and 3. Plots (b) and (c) show axial location

%The centerline velocity decay and radial profiles show good agreement, as with the previous cases, with a slight improvement in the radial velocity profile. The vrms profiles are significantly improved along the centerline compared to TMA and TMB. The centerline anomaly is not completely removed, however, since implemented eddies are still subject to the geometric distortion of the TMB map, whose influence is felt by the τ−1 calculation for subsequent eddy events. e

\subsection{Jet Flame}
\label{sub:jetflame}

% TEXT COPIED FROM CYLINDRICAL ODT PAPER
%The newcylindrical ODTformulation is demonstrated in a round, reacting, turbulent jet flame.Results are compared to the experimental DLR-Aflame ofMeier et al. [19,36]. Heat release and temperature- and compositiondependent transport and thermodynamic properties are key characteristics of this flow. This canonical flame configuration has been used extensively to study and validate turbulent combustion models. Pitsch [41] studied differential diffusion effects in this flame using a classical unsteady flamelet model. Lindstedt and Ozarivsky [35] investigated it using a joint PDF approach, while Wang and Pope [52] used an LES/PDF model combined with a flamelet/progress-variable (FPV) model. Fairweather and Woolley [10] studied several chemical mechanisms using the DLR flame and a first-order conditional moment closure (CMC) model, and Lee and Choi [29,30] studied NO emissions using an Eulerian particle flamelet model. These studies consisted of computational fluid dynamics (CFD) simulations with subgrid modeling of the turbulent combustion process. In contrast, ODT is a low-dimensional representation of the entire flow, not just the subgrid scales. Previous ODT studies of turbulent jet flames have employed the temporal planar formulation, so we expect that the model may improve with a spatial cylindrical formulation that more closely matches the experimental configuration.
%The DLR-A fuel stream is mixture of 22.1% CH4, 33.2% H2, and 44.7% N2 (by volume) that issues into dry air. The fuel stream exits a nozzle with an inner diameter of 8 mm at a mean exit velocity of 42.2 ms−1. The coflowing air stream issues from a nozzle of diameter 140 mm at a velocity of 0.3 ms−1. The reported Reynolds number is 15,200.
%The ODT simulation uses these diameters and the experimentally reported velocity profile at the jet exit. In the nonreacting case, a small velocity was added uniformly to the profile, no velocity addition is needed here because of the slow-moving coflow air stream present alongside the reacting jet. The ODT simulation includes the buoyant source term but not the streamwise pressure gradient in Eq. (25). The fuel was diluted with N2 in the experimental flame to minimize radiative heat losses, and radiation is ignored in the simulation. This flame has a low Reynolds number, and the combustion chemistry proceeds quickly. The ODT simulation transports the following species: O2, N2, CH4, H2, H2O, CO2. Reactions are assumed to proceed to products of complete combustion:
%H2 + 1/2O2 → H2O, CH4 + 2O2 → CO2 + 2H2O,
%(38) (39)
%and simple, fast reaction rates are applied. These assumptions are not reasonable for the DLR-A flame. The primary purpose of these simulations is to illustrate ODT in a reacting jet with variable properties and heat release.
%Z = 400. The values of C and βles were adjusted to give good agreement of the jet evolution with the experimental data, while the value of Z was taken as the same value as for the jet in Sect. 3.2. Results of ODT simulations were performed using the TMB triplet map with parameters C = 20, βles = 17, and 400 Th l f C d β djtd t i d t f th jt lti ith th
%the simulation are presented in Figs. 11 and 12. One thousand independent flow realizations were performed, and the results ensemble averaged. Downstream distance, y, and radial position, r, are normalized by the jet diameter D.

@@

@@
